### 2022-11-03 austins_analysis.rmd
First Commit Imported data, initialized Austin's .rmd and the main .rmd

# Analysis of the data by Austin
```{r}
# Import packages
library(tidyverse)
library(ggplot2)
```
```{r}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv")
```

```{r}
# Initial data exploration

# View the first 6 rows of the data
head(data)
# Available variables
names(data)
```