### 2022-11-03 main.rmd

# This r-markdown file is used to import the data and all of our own personal models as a final document