### 2022-11-12 main.rmd
Split Test and Training data with Group

# This r-markdown file is used to import the data and all of our own personal models as a final document

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
```

```{r}
set.seed(420)
data <- read.csv("breast_cancer_diagnostic_data.csv")
data <- subset(data, select = -c(X))
split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
train <- data[split == 0, ]
test <- data[split== 1, ]
#write.csv(train,"training.csv", row.names = FALSE)
#write.csv(test,"test.csv", row.names = FALSE)
```