### 2022-12-01 austins_analysis.rmd

# Analysis of the data by Austin
```{r}
# Import packages
library(tidyverse)
library(ggplot2)
```
```{r}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv")
```

```{r}
# Initial data exploration

# View the first 6 rows of the data
head(data)
# Available variables
names(data)
# Removing the ID column
data <- data[,-1]

```

```{r}
# Check for missing values
sum(is.na(data))
head(data[which(is.na(data)), ])
```

```{r}
# Looking for correlations between variables
library(GGally)
ggpairs(data, columns = 1:10)
ggpairs(data, columns = c(1, 11:20))
ggpairs(data, columns = c(1, 21:30))
```

Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst

```{r}
#logistic regression
data$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)
logistic_model <- glm(diagnosis_binary ~ radius_mean + perimeter_worst + area_worst + concave.points_worst, data = data, family = binomial)
summary(logistic_model)
```

```{r}
ggplot(data, aes(x = radius_mean, y = texture_worst, color=diagnosis)) + geom_point()
```

```{r}
```

```{r}

```