### 2022-12-01 paper.rmd

---
title: "Paper"
author: "Austin Lackey, Ethan Powers and Danny Laposata"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Sample Output
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
set.seed(420)
data <- read.csv("breast_cancer_diagnostic_data.csv")
data <- subset(data, select = -c(X))
split <- sample(c(rep(0, 0.6 * nrow(data)), rep(1, 0.4 * nrow(data))))
train <- data[split == 0, ]
test <- data[split== 1, ]
#write.csv(train,"training.csv", row.names = FALSE)
#write.csv(test,"test.csv", row.names = FALSE)
```

```{r}
print(names(train))
```