### 2022-12-02 austins_analysis.rmd
Logistic Regression Model

# Analysis of the data by Austin
```{r}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
```

```{r}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv")
training <- read.csv("training.csv")
test <- read.csv("test.csv")

# Data Cleaning
# Removing the ID column
data <- data[,-1]
test <- test[,-1]
training <- training[,-1]
# Removing the X column from data
data <- data[,1:(length(data)-1)]
# Check for missing values
print(paste(sum(is.na(data)), "missing values in the data"))
# Available variables
print(paste("There are", length(data), "variables available"))
print(names(data))
```

```{r}
# Looking for correlations between variables
g1 <- ggpairs(data, columns = 1:10)
g2 <- ggpairs(data, columns = c(1, 11:20))
g3 <- ggpairs(data, columns = c(1, 21:31))
```

Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst

```{r}
# Setting up a binary column for diagnosis
data$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)
training$diagnosis_binary <- ifelse(training$diagnosis == "M", 1, 0)
test$diagnosis_binary <- ifelse(test$diagnosis == "M", 1, 0)
```

```{r}
#logistic regression using predictors: radius_mean, area_worst, concave.points_worst
logistic_model <- glm(data = training, diagnosis_binary ~ radius_mean + area_worst + concave.points_worst, family = binomial)
print(summary(logistic_model))
logistic_model.prob <- predict(logistic_model, test, type = "response")
logistic_model.pred <- ifelse(logistic_model.prob > 0.5, 1, 0)
logistic_model.pos <- logistic_model.prob[logistic_model.pred == 1]
logistic_model.neg <- logistic_model.prob[logistic_model.pred == 0]
logistic_model_error <- test$diagnosis_binary != as.integer(logistic_model.pred)
ggplot() + geom_point(aes(x = test$radius_mean, y=test$perimeter_worst, color = test$diagnosis))
print(ggplot() + geom_point(aes(x = test$texture_mean, y=test$perimeter_mean, color = !logistic_model_error)) + labs(title = "Logistic Regression", x="texture_mean", y="perimeter_mean", color="Correct Prediction"))
```

```{r}
mean(logistic_model_error)
```

```{r}
ggplot(data, aes(x = radius_mean, y = texture_worst, color=diagnosis)) + geom_point()
```

```{r}
```

```{r}

```