### 2022-12-07 austins_analysis.rmd
updated analysis

# Analysis of the data by Austin
```{r package_load}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
```

```{r data_load}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv")
training <- read.csv("training.csv")
test <- read.csv("test.csv")

# Data Cleaning
# Removing the ID column
data <- data[,-1]
test <- test[,-1]
training <- training[,-1]
data$diagnosis <- as.factor(data$diagnosis)
test$diagnosis <- as.factor(test$diagnosis)
training$diagnosis <- as.factor(training$diagnosis)
# Removing the X column from data
data <- data[,1:(length(data)-1)]
# Check for missing values
print(paste(sum(is.na(data)), "missing values in the data"))
# Available variables
print(paste("There are", length(data), "variables available"))
print(names(data))
```

```{r ggpairs}
# Looking for correlations between variables
g1 <- ggpairs(data, columns = 1:10)
g2 <- ggpairs(data, columns = c(1, 11:20))
g3 <- ggpairs(data, columns = c(1, 21:31))
```

```{r ggpairs_plot}
# print(g1)
# print(g2)
# print(g3)
```

Variables to note: radius_mean, perimeter_worst, area_worst, concave.points_worst

```{r data_cleaning}
# Setting up a binary column for diagnosis
dataBinary <- data
trainingBinary <- training
testBinary <- test
dataBinary$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)
trainingBinary$diagnosis_binary <- ifelse(training$diagnosis == "M", 1, 0)
testBinary$diagnosis_binary <- ifelse(test$diagnosis == "M", 1, 0)
testBinary <- subset(testBinary, select = -diagnosis)
trainingBinary <- subset(trainingBinary, select = -diagnosis)
```

Logistic Regression

```{r logistic_regression}
# Logistic Regression without cross validation
logistic_regression <- glm(diagnosis_binary ~ ., data = trainingBinary, family = binomial)
logistic_pred <- predict(logistic_regression, testBinary, type="response")
logistic_pred <- ifelse(logistic_pred > 0.5, "M", "B")
logistic_pred <- data.frame(logistic_pred)$logistic_pred
logistic_confusion <- confusionMatrix(logistic_pred, as.character(test$diagnosis), positive = "M")
```

```{r train_setup}
train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
```

Random Forrest
```{r random_forrest}
#Random Forrest
random_forrest <- train(diagnosis ~ ., data = training, method = "ranger", metric="ROC", trControl = train_control, preProcess=c("center", "scale"))
rf_pred <- predict(random_forrest, test)
rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
print(rf_confusion)
```

```{r random_forrest_plot}
#Plotting the false positives and true positives for the random forest model
false_pred <- !(test$diagnosis != rf_pred)
false_negative <- ifelse((test$diagnosis == "M") & (rf_pred == "B"), "FalseNeg", "Okay")
ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Random Forrest Model False Negatives")
```

Random Forrest 2

```{r random_forrest_2}
x <- seq(1, 100, 1)
tuning_grid <- expand.grid(trees = x, rmse=NA)
for (i in seq_len(nrow(tuning_grid))) {
    # rf_fit <- train(diagnosis ~ ., 
    #     method = "ranger",
    #     metric="ROC",
    #     trControl = train_control,
    #     data = training, 
    #     num.trees = tuning_grid$trees[i])
    rf_fit <- randomForest(diagnosis ~ ., data = training, ntree = tuning_grid$trees[i])
    rf_pred <- predict(rf_fit, test)
    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
    tuning_grid$rmse[i] <- tibble(rf_confusion$table)[[2,1]][1]/sum(tibble(rf_confusion$table))
}
```

```{r random_forrest_2_plot}
ggplot() + geom_line(aes(x=x, y=tuning_grid$rmse))
```

KNN
```{r knn}
#KNN
knn <- train(diagnosis ~ ., data = training, method = "knn", metric="ROC", trControl = train_control)
knn_pred <- predict(knn, test)
knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
print(knn_confusion)
varsKNN <- rownames(data.frame(varImp(knn)[1]))[1:10]
```

```{r knn_plot}
print(ggplot(varImp(knn)) + labs(title = "KNN Variable Importance"))
```

```{r knn_plot2}
#Plotting the false positives and true positives for the knn model
false_pred <- !(test$diagnosis != knn_pred)
false_negative <- ifelse((test$diagnosis == "M") & (knn_pred == "B"), "FalseNeg", "Okay")
ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "KNN Model False Negatives")
```

Radial Support Vector Machine
```{r svm}
#Radial Support Vector Machine
svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
svm_pred <- predict(svm, test)
svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
print(svm_confusion)
varsSVM <- rownames(data.frame(varImp(svm)[1]))[1:10]
```

```{r svm_plot}
print(ggplot(varImp(svm)))
```

```{r svm_plot}
#Plotting the false positives and true positives for the logistic regression model
false_pred <- !(test$diagnosis != svm_pred)
false_negative <- ifelse((test$diagnosis == "M") & (svm_pred == "B"), "FalseNeg", "Okay")
ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "SVM Model False Negatives")
```

Logistic Regression
```{r logistic}
#Logistic Regression
logistic <- train(diagnosis ~ ., data = training, method = "glm", trControl = train_control, metric="ROC")
logistic_pred <- predict(logistic, test)
logistic_confusion <- confusionMatrix(logistic_pred, test$diagnosis, positive="M")
print(logistic_confusion)
logistic$finalModel
varsLOGISTIC <- rownames(data.frame(varImp(logistic)[1]))[1:10]
```

```{r}
#Plotting the false positives and true positives for the logistic regression model
false_pred <- !(test$diagnosis != logistic_pred)
false_negative <- ifelse((test$diagnosis == "M") & (logistic_pred == "B"), "FalseNeg", "Okay")
ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Logistic Regression Reduced Model False Negatives")
```

```{r logistic_plot}
print(ggplot(varImp(logistic)) + labs(title = "Logistic Variable Importance"))
```

```{r logistic_2}
logistic2 <- train(diagnosis ~ area_mean + area_worst + concave.points_mean + texture_mean + radius_worst + concavity_se + fractal_dimension_se + concavity_worst + concave.points_worst, data = training, method = "glm", trControl = train_control, metric="ROC")
logistic2_pred <- predict(logistic2, test)
logistic2_confusion <- confusionMatrix(logistic2_pred, test$diagnosis, positive="M")
print(logistic2_confusion)
```

```{r}
#Plotting the false positives and true positives for the logistic regression 2
false_pred <- !(test$diagnosis != logistic2_pred)
false_negative <- ifelse((test$diagnosis == "M") & (logistic2_pred == "B"), "FalseNeg", "Okay")
ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Logistic Regression Reduced Model False Negatives")
```

```{r logistic_2_plot}
```

```{r}
```