### 2022-12-08 ethan_temp_powerpoint.rmd
Add files via upload

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: "December 9th, 2022"
output: 
  powerpoint_presentation: 
    reference_doc: template.pptx

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
library(leaps)
library(caTools)
library(tidymodels)
library(glmnet)
library(pls)
library(forecast)
library(GGally)

set.seed(420)

dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
data <- dataTemp %>% select_if(~ !any(is.na(.)))
test <- read.csv("test.csv")
training <- read.csv("training.csv")
```

## Background

- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features
- Reduce Type II Error (False-Negatives)
- Feature Selection

## Data

- Data collected from "Breast Cancer Wisconsin (Diagnostic) Data Set" from Kaggle
- Response: Diagnosis (Benign/Malignant)
- Predictors: 10 variables divided between  Mean, SE, and "Worst"
- Training/Test Split (60/40)

## EDA

- Different measurements of each variable are correlated

- Relationship with Diagnosis visible for each measurement

```{r, fig.height=3, fig.width=4, fig.align='right', fig.pos='H'}
#source("Danny's Analysis.R", echo = F, local = T)

#train[,c(1,2,12,22)] |>
#  ggpairs(title = "Correlation b/w Radius Measurements")
```

## Classification Analysis: Austin

## Classification Analysis: Danny

## Regression Analysis: Ethan
```{r}
dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
testTot <- read.csv("test.csv")
trainingTot <- read.csv("training.csv")

dataE <- data
corr_df <- dataE[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
ggpairs(corr_df)
```
-Test

## Model Assesments

## Works Cited

- Breast Cancer Wisconsin (Diagnostic) Data Set: https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount