### 2022-12-08 austins_temp_powerpoint.rmd
updated

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: "December 9th, 2022"
output: 
  powerpoint_presentation: 
    reference_doc: template.pptx

---
```{r setup, include=FALSE}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
library(class)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
set.seed(420)
```

```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
training <- read.csv("training.csv") # Training Data
test <- read.csv("test.csv")  # Test Data
# Data Cleaning
data <- data[,-1] # Remove ID column
test <- test[,-1] # Remove ID column
training <- training[,-1] # Remove ID column
# Convert diagnosis to factor
data$diagnosis <- as.factor(data$diagnosis)
test$diagnosis <- as.factor(test$diagnosis)
training$diagnosis <- as.factor(training$diagnosis)
```

## Background

- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features (Malignant or Benign)
- Reduce Type II Error (False-Negatives)
  - Feature and Model Selection

# Models

```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
```

## KNN

```{r austin_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# KNN
knn <- train(diagnosis ~ ., data = training, method = "knn", metric="ROC", trControl = train_control)
knn_pred <- predict(knn, test)
knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
varsKNN <- rownames(data.frame(varImp(knn)[1]))[1:10]
```

```{r austin_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(knn_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "KNN Confusion Matrix", x = "Reference", y = "Prediction") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=2+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Random Forrest

```{r austin_random_forrest, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Logistic Regression without cross validation
random_forrest <- train(diagnosis ~ ., data = training, method = "ranger", metric="ROC", trControl = train_control, preProcess=c("center", "scale"))
rf_pred <- predict(random_forrest, test)
rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
```

```{r austin_random_forrest_plot, echo=FALSE, message=FALSE, warning=FALSE}
# Plot confusion matrix
ggplot(data.frame(rf_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Random Forrest Confusion Matrix", x = "Reference", y = "Prediction") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=2+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

```{r austin_random_forrest_plot2, echo=FALSE, message=FALSE, warning=FALSE}
#Plotting the false positives and true positives for the logistic regression model
false_pred <- !(test$diagnosis != rf_pred)
false_negative <- ifelse((test$diagnosis == "M") & (rf_pred == "B"), "FalseNeg", "Okay")
ggplot() + geom_point(aes(x=test$texture_mean, y=test$radius_mean, color=false_negative), size=2) + labs(title = "Random Forrest Model False Negatives")
```

## Radial Support Vector Machine

```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Radial Support Vector Machine
svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
svm_pred <- predict(svm, test)
svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")
```

```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data.frame(svm_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Radial SVM Confusion Matrix", x = "Reference", y = "Prediction") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=2+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
```

```{r austin_svm22, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
svm2_pred <- predict(svm2, test)
svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")
```

```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(svm2_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Radial SVM Confusion Matrix", x = "Reference", y = "Prediction") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=2+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

```{r}
learn_rf <- randomForest(diagnosis~., data=training, ntree=500, proximity=T, importance=T)
```

```{r}
acc_test <- numeric() 
for(i in 1:30){
    predict <- knn(train=training[,-1], test=test[,-1], cl=training$diagnosis, k=i, prob=T)
    acc_test <- c(acc_test,mean(predict==test[,1]))
}
```