### 2022-12-09 powerpoint.rmd

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: "December 9th, 2022"
output: 
  powerpoint_presentation: 
    reference_doc: template.pptx

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 14, fig.height = 8)
```

## Background

- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features
- Reduce Type II Error (False-Negatives)
- Feature Selection

## Data

- Data collected from Kaggle
- Response: Diagnosis (Benign/Malignant)
- Mean, SE, and "Worst" measurements of 10 variables
- Training/Test Split (60/40)

## Features

- 
-

## EDA

```{r}
ggplot(train) +
  geom_point(aes(x = concavity_mean, y = radius_mean, col = diagnosis))+
  labs(x = "Concavity Mean", y = "Radius Mean", title = "Severity of Concavity vs. Radius of Tumor")

```

Different measurements of each variable are correlated

Relationship with Diagnosis visible for each measurement

## Classification Analysis: Austin

## Classification Analysis: Danny

```{r}
### Packages

library(tidyverse)
library(GGally)
library(ggplot2)
library(tidyr)
library(tidymodels)
library(vip)
library(caTools)
library(xgboost)
library(discrim)
library(glmnet)
library(kableExtra)

### Gathering Data

train <- read.csv("training.csv")
test <- read.csv("test.csv")

train |>
  select(-id) -> train

test |>
  select(-id) -> test

train$diagnosis <- as.factor(train$diagnosis)
test$diagnosis <- as.factor(test$diagnosis)
```

- Logistic Regression
- K-Nearest Neighbors
- QDA
- Random Forest
- xGBoost
- Support Vector Classifier
- Lower prediction cutoff to account for Type II Error

## K-Nearest Neigbors Model

```{r}
set.seed(445)
### KNN

cancer_10foldcv <- vfold_cv(train, v = 10)
neighbors_df <- data.frame(neighbors = c(1,3,5,7,10))

knn_spec <- the_model <- nearest_neighbor(neighbors = tune("neighbors")) %>%
  set_engine("kknn") %>% 
  set_mode("classification")

knn_rec <- recipe(diagnosis ~ ., data = train)

workflow() |>
  add_model(knn_spec) |>
  add_recipe(knn_rec) -> knn_wf

knn_wf |>
  tune_grid(resamples = cancer_10foldcv, grid = neighbors_df) -> knn_tune

knn_final <- finalize_workflow(knn_wf, select_best(knn_tune, metric = "accuracy"))
knn_final_fit <- fit(knn_final, data = train)

knn_final_fit |>
  augment(new_data = test) -> knn_test_res

knn_test_res |>
  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> knn.cm

knn_test_res |>
  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
  mutate(error = 1 - .estimate) |>
  pull(error) -> knn.err

100 - (round(knn.err, 4) * 100) -> knn.err

neighbors_best <- select_best(knn_tune, metric = "accuracy") |> pull(neighbors)
```

- Tuned neighbors value to `r neighbors_best`
- 2 False-Negatives
- `knn.err` % prediction accuracy

```{r}
ggplot(data.frame(knn.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "K-Nearest Neighbors Confusion Matrix", x = "Prediction", y = "Truth") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +
  guides(fill = F)
```

## Support Vector Classifier Model

```{r}
set.seed(445)
### SVC

df_cost <- grid_regular(cost(), levels = 10)

svm_linear_tune_spec <- svm_poly(degree = 1, cost = tune("cost")) %>%
  set_mode("classification") %>%
  set_engine("kernlab", scaled = FALSE)

svm_linear_rec <- recipe(diagnosis ~ ., data = train)

svm_linear_wf <- workflow() |>
  add_model(svm_linear_tune_spec) |>
  add_recipe(svm_linear_rec)

tune_fit <- svm_linear_wf |>
  tune_grid(resamples = cancer_10foldcv, grid = df_cost)

svm_linear_final <- finalize_workflow(svm_linear_wf, select_best(tune_fit, metric = "accuracy"))

svm_linear_final |>
  fit(data = train) -> svm_linear_final_fit

svm_linear_final_fit |>
  extract_fit_engine() -> svm_linear_final_fit_engine

svm_linear_final_fit |>
  augment(new_data = test) |>
  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> svm.cm

svm_linear_final_fit |>
  augment(new_data = test) |>
  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
  mutate(error = 1 - .estimate) |>
  pull(error) -> svm.err

100 - (round(svm.err, 4) * 100) -> svm.err

cost_best <- select_best(tune_fit, metric = "accuracy") |> pull(cost)
```

- Tuned cost value to `r cost_best`
- 2 False-Negatives
- `svm.err` % prediction accuracy

```{r}
ggplot(data.frame(svm.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Support Vector Classifier Confusion Matrix", x = "Prediction", y = "Truth") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
  guides(fill = F)
```

## Random Forest Model

```{r}
### Random Forest

rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
  set_engine("randomForest", importance = TRUE) |>
  set_mode("classification")

rf_fit <- rf_spec |>
  fit(diagnosis ~ ., data = train)

rf_fit |>
  augment(new_data = test) |>
  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
  conf_mat(truth = diagnosis, estimate = pred_lower_cutoff) -> rf.cm

rf_fit |>
  augment(new_data = test) |>
  mutate(pred_lower_cutoff = factor(ifelse(.pred_M > 0.25, "M", "B"))) |>
  accuracy(truth = diagnosis, estimate = pred_lower_cutoff) |>
  mutate(error = 1 - .estimate) |>
  pull(error) -> rf.err

100 - (round(rf.err, 4) * 100) -> rf.err
```

- 1 False-Negative
- `rf.err` % prediction accuracy

```{r, echo = F}
ggplot(data.frame(rf.cm$table), aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Random Forest Confusion Matrix", x = "Prediction", y = "Truth") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798") +  
  guides(fill = F)

vip(rf_fit)
```

## Regression Analysis: Ethan

## Model Assesments

## Works Cited

- Insert link to kaggle dataset here