### 2022-12-09 austins_temp_powerpoint.rmd

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: "December 9th, 2022"
output: 
  powerpoint_presentation: 
    reference_doc: template.pptx

---
```{r setup, include=FALSE}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
library(class)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
set.seed(420)
```

```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
training <- read.csv("training.csv") # Training Data
test <- read.csv("test.csv")  # Test Data
# Data Cleaning
data <- data[,-1] # Remove ID column
test <- test[,-1] # Remove ID column
training <- training[,-1] # Remove ID column
# Convert diagnosis to factor
data$diagnosis <- as.factor(data$diagnosis)
test$diagnosis <- as.factor(test$diagnosis)
training$diagnosis <- as.factor(training$diagnosis)
```

## Background

- Many people are diagnosed with breast cancer each year
- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features (Malignant or Benign)
- *Modivation:* Reduce Type II Error (False-Negatives)
  - Feature and Model Selection

# Models

```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
```

# KNN Model

## Tuning KNN

```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)

for (i in 1:nrow(type_two_error_knn)) {
    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
    knn_pred <- predict(knn, test)
    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
}
print(which.min(type_two_error_knn$typetwoerror))
knn1_k <- which.min(type_two_error_knn$typetwoerror)
knn_tte <- min(type_two_error_knn$typetwoerror)
```

```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
```

## Tuned KNN

```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(420)
tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
knn_pred <- predict(tuned_knn, test)
knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
```

```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "KNN Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Tuned KNN Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- $k=$ `r knn1_k` neighbors
- Type II Error: `r knn_tte`
- Accuracy: `r knn_confusion$overall[1]`

# Random Forrest

## Tuning Random Forrest

```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)

for (i in 1:nrow(type_two_error_rf)) {
    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
    rf_pred <- predict(rf, test)
    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
}
print(which.min(type_two_error_rf$typetwoerror))
rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
rf_tte <- min(type_two_error_rf$typetwoerror)
```

```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
```

## Tuned Random Forrest

```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
rf_pred_tuned <- predict(tuned_rf, test)
rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
```

```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Tuned Random Forrest Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- $n=$ `r rf1_ntree` trees
- Type II Error: `r rf_tte`
- Accuracy: `r rf_confusion$overall[1]`

## Radial Support Vector Machine

```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Radial Support Vector Machine
svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
svm_pred <- predict(svm, test)
svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")

svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
```

```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Radial Support Vector Machine Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- Type II Error: `r svm_tte`
- Accuracy: `r svm_confusion$overall[1]`

## Tuned Radial Support Vector Machine

```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
svm2_pred <- predict(svm2, test)
svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")

svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)

```

```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Tuned Radial Support Vector Machine Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- Type II Error: `r svm2_tte`
- Accuracy: `r svm2_confusion$overall[1]`