### 2022-12-10 austin_temp_paper.rmd

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: AUSTIN'S TEMPLATE
date: "December 9th, 2022"
output: 
  pdf_document:
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    df_print: paged
    fig_caption: TRUE
    highlight: tango
---
```{r setup, include=FALSE}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
library(class)
set.seed(420)
```

# Background (Should we do an Abstract?)

*Insert text here*

# Data

*Insert text here*

## Variable Descriptions

*Insert text here*

# Explatory Data Analysis

*Insert text here*

# Austin's Classification Proccess and Analysis

*Insert text here*

# Dannys's Classification Proccess and Analysis

*Insert text here*

# Ethans's Regression Proccess and Analysis

*Insert text here*

# Analysis Summary

*Insert text here*

# Potential Improvements

*Insert text here*

# Works Cited

- Breast Cancer Wisconsin (Diagnostic) Data Set:
    - https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
- Definition of Features (Variables):
    - https://www.causeweb.org/usproc/sites/default/files/usclap/2017-2/Evaluating_Benign_and_Malignant_Breast_Cancer_Cells_from_Fine-Needle_Aspirates.pdf