### 2022-12-15 ethans_temp_paper.rmd
Add files via upload
Ethan part of paper

---
title: "Breast Cancer Diagnosis Classification Analysis"
author: ETHANS'S TEMPLATE
date: "December 9th, 2022"
output: 
  pdf_document:
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    df_print: paged
    fig_caption: TRUE
    highlight: tango
    extra_dependencies: ["float"]

---
```{r setup, include=FALSE}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
library(class)
#
library(dplyr)
library(leaps)
library(caTools)
library(tidymodels)
library(glmnet)
library(forecast)
library(plyr)
library(readr)
library(repr)
library(kableExtra)

set.seed(420)

#
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#tinytex::install_tinytex()

```

# Background (Should we do an Abstract?)

*Insert text here*

# Data

*Insert text here*

## Variable Descriptions

After removing the `ID` column, we were left with `31 variables`. 
For this project our response variable is `diagnosis` as this variable tells whether an observation is Malignant (cancerous cells): `M` or Benign (noncancerous cells): `B`.
The predictor variables make up the remaining 30 columns; however there are only actually 10 variables that each have 3 measurements.
These measurements are `mean`, `standard error (SE)`, and `worst` (the mean of the 3 largest values of a variables).

Below is a brief explanation of what each of the 10 variables represent:

- Radius: Average Distance from cell center to cell perimeter
- Texture: Standard deviation of gray-scale values; brightness of pixel of cell
- Perimeter: Distance around nucleus boundary
- Area: Area of the nucleus
- Smoothness: Variation in cell's radial lengths
- Compactness: The Perimeter^2/Area
- Concavity: Size of the indention in nucleus boundary
- Concave Points: Number of points on indented section of nucleus boundary
- Symmetry: Deviation of the nuclei shape from the ideal measurement
- Fractal Dimension: Measurement of irregularity in nucleus boundary

# Explatory Data Analysis

*Insert text here*

# Austin's Classification Proccess and Analysis

*Insert text here*

# Dannys's Classification Proccess and Analysis

*Insert text here*

# Ethans's Regression Proccess and Analysis

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}

dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]

testtemp <- read.csv("test.csv")
testTot <- testtemp[,!names(testtemp) %in% c("id")]

traintemp <- read.csv("training.csv")
trainTot <- traintemp[,!names(traintemp) %in% c("id")]
```

While Austin and Danny did Classification Analysis on `Type-II error rate`, I decided to do a Regression Analysis using the variable importance from their analysis's to compute the `training` and `test` error rate of those variables. 
This is to see how much error exists in a variable that is important in calculating if a cell is Malignant or Benign.
However, I will be focused on `test` error in my results as its more important to see how our results will be on future data rather than what we have already trained. The regression models that I will be using are:

    - Linear Regression
    - Ridge Regression
    - Lasso Regression

## Variables

In order to find the variables to run the Regression Analysis on, I had Austin and Danny run a variable importance plot on the data they collected and took the top 5 variables from each of their results, which are listed below.

- Austin Top 5 Importance Variables:

  1. *`area_mean`
  2. *`area_worst`
  3. *`concave.points_mean`
  4. `texture_mean`
  5. *`radius_worst`

- Danny Top 5 Importance Variables

  1. `concave.points_worst`
  2. *`concave.points_mean`
  3. *`area_worst`
  4. *`area_mean`
  5. *`radius_worst`

After I got both Austin's and Danny's top 5 importance variables, I noticed that there were 4 common variables in each of their results (designated with a '*' in the results above).
I then decided to see how the 4 common importance variables correlate to one another which can be seen in the pairwise scatterplot below:

```{r correlation_plot, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
corr_df <- dataTot[,c( 'area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
ggpairs(corr_df) + labs(title = "Figure 1: Pairwise Scatterplot of Common Importance Variables")
```

What I found interesting was that `area_worst` & `raduis_worst` were more correlated that `area_worst` & `area_mean`. What I initially thought before making `Figure 1` was that `area_worst` & `area_mean` would be more correlated than the other 2 variables as they are both under the same variable category.

## area_mean

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Train
am_lm <- lm(area_mean ~ ., data = trainTot)
pred_amtrain<- predict(am_lm, trainTot)
amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
#amtrain_log
```
```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Test
pred_amtest <- predict(am_lm, testTot)
amtest_log <- postResample(pred_amtest, testTot$area_mean)
#amtest_log
#summary(am_lm)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
dummy <- dummyVars(area_mean~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$area_mean

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$area_mean

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min

eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquared = R_square)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
amr_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
amr_test <- eval_results(y_test, predictions_test, testTot)

#print(amr_train)
#print(amr_test)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
lambdas <- 10^seq(2, -3, by = -.1)
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lambda_best <- lasso_reg$lambda.min 
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
aml_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
aml_test <- eval_results(y_test, predictions_test, testTot)

#print(aml_train)
#print(aml_test)
```

For `area_mean` after running the 3 regression models, I found the lowest `test error` is from the ridge regression model with an `RMSE` of `23.70795` with a `Rsquared` of `99.489%`. 
This mean that using ridge regression, we can explain `99.489%` of the variance in `area_mean` in Austin's and Danny's models.
The ridge regression corresponding `training error` is `17.87197` with a `Rsquared` of 	`99.759%`.
This corresponds to that `99.759%` of our variance in `area_mean` was trained in Austin's and Danny's models.

```{r area_mean_table, echo=FALSE, message=FALSE, warning=FALSE}
table_am <- as_tibble(rbind(amtrain_log,
      amtest_log,
      amr_train,
      amr_test,
      aml_train,
      aml_test)) %>%
    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
    select(model, RMSE, Rsquared)
table_am %>% 
  kbl(digits = 5 ) %>% 
  kable_styling(bootstrap_options = "striped") %>%
  kable_styling(latex_options = "HOLD_position")
```

## Area_worst

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#Test
aw_lm <- lm(area_worst ~ ., data = trainTot)
pred_awtest <- predict(aw_lm, testTot)
awtest_log <- postResample(pred_awtest, testTot$area_worst)
#awtest_log
#summary(am_lm)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#Train
pred_awtrain<- predict(aw_lm, trainTot)
awtrain_log <- postResample(pred_awtrain, trainTot$area_worst)
#awtrain_log
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
dummy <- dummyVars(area_worst~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$area_worst

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$area_worst

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min

eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquared = R_square)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
awr_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
awr_test <- eval_results(y_test, predictions_test, testTot)

#print(awr_train)
#print(awr_test)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
lambdas <- 10^seq(2, -3, by = -.1)
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lambda_best <- lasso_reg$lambda.min 
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
awl_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
awl_test <- eval_results(y_test, predictions_test, testTot)

#print(awl_train)
#print(awl_test)
```

In `area_worst` after running the 3 regression models, I found the lowest `test error` is from the ridge regression model with an `RMSE` of `42.28132` with a `Rsquared` of `99.288%`. 
This mean that using ridge regression, we can explain `99.288%` of the variance in `area_worst` in Austin's and Danny's models.
The ridge regression corresponding `training error` is `27.36780` with a `Rsquared` of `99.799%`.
This corresponds to that `99.799%` of our variance in `area_worst` was trained in Austin's and Danny's models.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_aw <- as_tibble(rbind(awtrain_log,
      awtest_log,
      awr_train,
      awr_test,
      awl_train,
      awl_test)) %>%
    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
    select(model, RMSE, Rsquared)
table_aw %>% 
  kbl(digits = 5 ) %>% 
  kable_styling(bootstrap_options = "striped") %>%
  kable_styling(latex_options = "HOLD_position")
```

## concave.points_mean

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#Train
cvptsm_lm <- lm(concave.points_mean ~ ., data = trainTot)
pred_cvptsm_train<- predict(cvptsm_lm, trainTot)
cvptsm_train_log <- postResample(pred_cvptsm_train, trainTot$concave.points_mean)
#cvptsm_train_log
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#Test
pred_cvptsm_test <- predict(cvptsm_lm, testTot)
cvptsm_test_log <- postResample(pred_cvptsm_test, testTot$concave.points_mean)
#cvptsm_test_log
#summary(am_lm)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
dummy <- dummyVars(concave.points_mean~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$concave.points_mean

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$concave.points_mean

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min

eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquared = R_square)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
cvptsmr_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
cvptsmr_test <- eval_results(y_test, predictions_test, testTot)

#print(cvptsmr_train)
#print(cvptsmr_test)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
lambdas <- 10^seq(2, -3, by = -.1)
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lambda_best <- lasso_reg$lambda.min 
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
cvptsml_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
cvptsml_test <- eval_results(y_test, predictions_test, testTot)

#print(cvptsml_train)
#print(cvptsml_test)
```

With `concave.points_mean` after running the 3 regression models, I found the lowest `test error` is from the linear regression model with an `RMSE` of `0.00574` with a `Rsquared` of `97.847%`. 
This mean that using ridge regression, we can explain `97.847%` of the variance in `concave.points_mean` in Austin's and Danny's models.
The ridge regression corresponding `training error` is `0.00467` with a `Rsquared` of `98.532%`.
This corresponds to that `98.532%` of our variance in `concave.points_mean` was trained in Austin's and Danny's models.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_am <- as_tibble(rbind(cvptsm_train_log,
      cvptsm_test_log,
      cvptsmr_train,
      cvptsmr_test,
      cvptsml_train,
      cvptsml_test)) %>%
    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
    select(model, RMSE, Rsquared)
table_am %>% 
  kbl(digits = 5 ) %>% 
  kable_styling(bootstrap_options = "striped") %>%
  kable_styling(latex_options = "HOLD_position")
```

## radius_worst

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#Train
rw_lm <- lm(radius_worst ~ ., data = trainTot)
pred_rw_train<- predict(rw_lm, trainTot)
rw_train_log <- postResample(pred_rw_train, trainTot$radius_worst)
#rw_train_log
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#Test
pred_rw_test <- predict(rw_lm, testTot)
rw_test_log <- postResample(pred_rw_test, testTot$radius_worst)
#rw_test_log
#summary(am_lm)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
dummy <- dummyVars(radius_worst~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$radius_worst

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$radius_worst

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min

eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquared = R_square)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
rwr_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
rwr_test <- eval_results(y_test, predictions_test, testTot)

#print(rwr_train)
#print(rwr_test)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
lambdas <- 10^seq(2, -3, by = -.1)
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lambda_best <- lasso_reg$lambda.min 
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
rwl_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
rwl_test <- eval_results(y_test, predictions_test, testTot)

#print(rwl_train)
#print(rwl_test)
```

For `radius_worst` after running the 3 regression models, I found the lowest `test error` is from the lasso regression model with an `RMSE` of `0.24120	` with a `Rsquared` of `99.713%`. 
This mean that using ridge regression, we can explain `99.713%` of the variance in `radius_worst` in Austin's and Danny's models.
The ridge regression corresponding `training error` is `0.15907` with a `Rsquared` of `99.900%`.
This corresponds to that `99.900%` of our variance in `radius_worst` was trained in Austin's and Danny's models.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
table_am <- as_tibble(rbind(rw_train_log,
      rw_test_log,
      rwr_train,
      rwr_test,
      rwl_train,
      rwl_test)) %>%
    mutate(model = c('Linear Train Error',  'Linear Test Error','Ridge Train Error', 'Ridge Test Error', 'Lasso Train Error', 'Lasso Test Error')) %>%
    select(model, RMSE, Rsquared)
table_am %>% 
  kbl(digits = 5 ) %>% 
  kable_styling(bootstrap_options = "striped") %>%
  kable_styling(latex_options = "HOLD_position")
```

# Analysis Summary

*Insert text here*

# Potential Improvements

*Insert text here*

# Works Cited

- Breast Cancer Wisconsin (Diagnostic) Data Set:
    - https://www.kaggle.com/datasets/uciml/breast-cancer-wisconsin-data?datasetId=180&sortBy=voteCount
- Definition of Features (Variables):
    - https://www.causeweb.org/usproc/sites/default/files/usclap/2017-2/Evaluating_Benign_and_Malignant_Breast_Cancer_Cells_from_Fine-Needle_Aspirates.pdf